---
title: "EDS 222: Final Project"
subtitle: "Mangrove Blue Carbon"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)

```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")

```


## Introduccion:

Question and why is important:

```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","monitoring_siteA_data_javier.csv")) |> clean_names()

mangrove_df_b <- read_csv(here::here("data","monitoring_siteB_data_javier.csv")) |> clean_names()


```

```{r}
#Wranggle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)

```

```{r}
summary(mangrove_df_b)
names(mangrove_df_b)
```


```{r}
#Changing some data
mangrove_df$height_cm[1253] <- 112
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)

```

Create Plots to visualize the data
```{r}
#Plotting
ggplot(mangrove_df, aes(x=plot_size_m2, y = total_tree_mg_c_ha, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Sample Plots - Site A",
       subtitle = "Tree kg C vs Height") +
    theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.background = element_rect(fill = "gray98", size = 4, colour = "gray92"),
    axis.text.x = element_text(angle=-45, hjust = 0),
    panel.background = element_rect(fill = "gray91"),
    panel.grid.major = element_line(colour = "grey70", size = 0.2),
    panel.grid.minor = element_line(color = "gray88"))


```
Create plots that are relevant to your questions

Carbon Sequastration (kg/tree) vs Size
(_ cd_chatting_m, y = total_tree_kg_c _)
```{r}
#Plotting variables of interest
ggplot(mangrove_df, aes(x=cd_chatting_m, y = total_tree_kg_c, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Carbon Sequastration (kg/tree)  vs Size")
```

Create a table that is grouped by the each plantation year and contains the total number of sample plots and the total number of samples trees per plantation.

```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n())
          

knitr::kable(mangrove_summary)

```

Create a table that shows the samples per each plot

```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot) |> 
  summarise(plot = unique(plot),
            sample_count = n())  

knitr::kable(plots)


```

Create a table of samples per plot but grouping by sample size

```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_m = mean(cd_chatting_m),
            mean_tot_carbon_kg_c = mean(total_tree_kg_c))

kable(mangrove_plots_summary)

```

Lets focus on the 2017 plantation year

```{r}
# Potential need to rename the plots as "Sample Plot#1", "Sample plot #2", etc
# Potential need to rename the plot size as "25m2", "77m2", "154m2"

mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  select(plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha)

kable(mangrove_2017)


```

```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n())

kable(plots_2017)

```

```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")
```

-   Create a table and calculate the mean and sd on total ca per hectare for each plot size

```{r}
#Statistical Table with interesting variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            plot_count = length(unique(plot)),
            mean_chatting_size_m = mean(cd_chatting_m),
            mean_tree_c_kg = mean(total_tree_kg_c),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_chatting_size_m = sd(cd_chatting_m),
            sd_tree_c_kg = mean(total_tree_kg_c),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2")) |> 
  mutate(plot_size_m2 = factor(plot_size_m2, levels = plot_size_m2)) 

kable(mangrove_plots_2017)
```

```{r}


ggplot(mangrove_plots_2017, aes(x=plot_size_m2 , y=mean_chatting_size_m)) +
  geom_col() +
  labs(title = "Tree Size means")

ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c_kg)) +
  geom_col() +
  labs(title = "Carbon mean per tree")


ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c_ha_mg)) +
  geom_col() +
  labs(title = "Carbon mean per tree per ha")

```


Option 1 For variable Selection. 
Total Carbon per tree for Hectare vs Sample Plot Type (Plot Type 25m2 vs 154m2)

```{r}

mangrove_2017 |> 
  filter(plot_size_m2 == 25.133 ) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram() 
```


```{r}
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938 ) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram() 
```


```{r}
#Variable Total Carbon per tree (mg)
mangrove_2017 |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha ,color = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type")
```


State your null and alternative hypotheses

**Response: For the null hypothesis you state that there is no difference for the carbon stored in each tree per hectare, between the sample plot size 25m2 and the sample size of 154m2.**

$$H_0: \mu_{154m2} - \mu_{25m2} = 0$$
 Calculate the Point Estimate
```{r}
#Calculate Point Estimate
names(mangrove_plots_2017)
point_est_1 = ( - mangrove_plots_2017$mean_tree_c_ha_mg[3] + mangrove_plots_2017$mean_tree_c_ha_mg[1])
print(point_est_1)

```

Calculate the Standard Error
```{r}
#Define the Standard Error
n_25m = mangrove_plots_2017[[1,2]]
n_154m = mangrove_plots_2017[[3,2]]

s_25m =  mangrove_plots_2017[[1,9]]
s_154m =  mangrove_plots_2017[[3,9]]

SE_1 = as.numeric(sqrt(s_25m^2/n_25m + s_154m^2/n_154m))
print(SE_1)


```

Calculate the Z-Score
```{r}

z_score_1 <- (point_est_1 - 0) / SE_1
z_score_1
```

Calculate the P-Value
```{r}
option1_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2==25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

```
**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 25m2 to the sample size of 154m2**

Check differences between manual and t.test automatic

```{r}
print(option1_ttest$p.value)

pval_1 = 2 * pnorm(point_est_1, mean = 0, sd = SE_1, lower.tail = F)
print(pval_1)

```



```{r}

crit_val_1 = qnorm(0.025, lower.tail = F)
ci_lower_1 = round(point_est_1 - crit_val_1*SE_1, 2)
ci_upper_1 = round(point_est_1 + crit_val_1*SE_1, 2)


print(paste0("There is 95% probability that between [", ci_lower_1,",",ci_upper_1, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 154m2"))
```


Option 2 For variable Selection
Total Carbon per tree vs Tree Size
```{r}
#Variable Total Carbon per tree (kg)
mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_line(alpha = 0.3, color = "deepskyblue4") +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size")
```

```{r}
mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c, col = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.2) +
  labs(title = "Tree kg C vs Size") +
  facet_wrap(~ plot_size_m2)

```

Starting the Statistical analysis

```{r}
#Calculate the lm of total carbon per tree in relation to their size
names(mangrove_2017)

lm_carbon <- lm(total_tree_kg_c ~ cd_chatting_m, data=mangrove_2017) |> 
  summary()
  
lm_carbon

```

Calculate Point Estimate, Standard Error, Point Estimate, Confident Intervals
```{r}

crit_val_95 <- qnorm(0.025, lower.tail = FALSE)
PE_carbon <- lm_carbon$coefficients[,"Estimate"][2]
SE_carbon <- lm_carbon$coefficients[,"Std. Error"][2]

ci_lower = round(PE_carbon - crit_val_95 * SE_carbon, 2)
ci_upper = round(PE_carbon + crit_val_95 * SE_carbon, 2)

ci_lower
ci_upper

```


**Confident Intervals: We are 95% confident that our point estimate lays within the confidence interval 2.5- 2.6 of  tree size.**
