---
title: "EDS 222: Final Project"
subtitle: "Mangrove Blue Carbon"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
---

## Introduccion:

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)

```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")

```

```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","monitoring_siteA_data_javier.csv")) |> clean_names()

mangrove_df_b <- read_csv(here::here("data","monitoring_siteB_data_javier.csv")) |> clean_names()


```

```{r}
#Wranggle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)


```

```{r}
#Changing some data
mangrove_df$height_cm[1253] <- 112
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)

```

```{r}
#Plotting
ggplot(mangrove_df, aes(x=plot_size_m2, y = total_tree_mg_c_ha, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Sample Plots - Site A",
       subtitle = "Tree kg C vs Height") +
    theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.background = element_rect(fill = "gray98", size = 4, colour = "gray92"),
    axis.text.x = element_text(angle=-45, hjust = 0),
    panel.background = element_rect(fill = "gray91"),
    panel.grid.major = element_line(colour = "grey70", size = 0.2),
    panel.grid.minor = element_line(color = "gray88"))


```

```{r}
#Ploting variables of interest
ggplot(mangrove_df, aes(x=cd_chatting_m, y = total_tree_kg_c, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Carbon Sequastration (kg/tree)  vs Size")
```

Create a table that is grouped by the each plantation year and contains the total number of sample plots and the total number of samples trees per plantation.

```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n())
          

knitr::kable(mangrove_summary)

```

Create a table that shows the samples per each plot

```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot) |> 
  summarise(plot = unique(plot),
            sample_count = n())  

knitr::kable(plots)


```

Create a table of samples per plot but grouping by sample size

```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_m = mean(cd_chatting_m),
            mean_tot_carbon_kg_c = mean(total_tree_kg_c))

kable(mangrove_plots_summary)

```

Lets focus on the 2017 plantation year

```{r}
mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  select(plot, plot_size_m2, cd_chatting_m,total_tree_kg_ctotal_tree_mg_c_ha)

kable(mangrove_2017)

```

```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n())

kable(plots_2017)

```

```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")
```


```{r}
#Statistical Table with interesing variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            plot_count = length(unique(plot)),
            mean_chatting_size_m = mean(cd_chatting_m),
            mean_tree_c_kg = mean(total_tree_kg_c),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_chatting_size_m = sd(cd_chatting_m),
            sd_tree_c_kg = mean(total_tree_kg_c),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

kable(mangrove_plots_2017)

```


Option 1 For variable Selection. 
Total Carbon per tree vs Tree Size
```{r}
#Variable Total Carbon per tree (kg)
mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.7, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size")
```

```{r}
mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c, col = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree kg C vs Size") +
  facet_wrap(~ plot_size_m2)

```

Option 2 For variable Selection

Total Carbon per tree for Hectare vs Sample Plot Type
```{r}
#Variable Total Carbon per tree (mg)
mangrove_2017 |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha ,color = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type")
```


Starting the Statistical analysis

```{r}
#Histogram of Total Carbon per Tree/Ha
ggplot(mangrove_2017, aes(x=total_tree_kg_c)) +
  geom_histogram() +
  facet_wrap(~plot_size_m2)
```

```{r}

ggplot(mangrove_2017, aes(x=total_tree_mg_c_ha)) +
  geom_histogram() + 
  facet_wrap(~plot_size_m2)#, ncol = 1)
```


-   Create a table and calculate the mean and sd on total ca per hectar for each plot size

 Plot Type 25m2 vs 154m2
```{r}
#Calulate Point Estimate
names(mangrove_plots_2017)
point_est_1 = (mangrove_plots_2017$mean_tree_c_ha_mg[3] - stat_plots_2017$mean_tree_c_ha_mg[1])
print(point_est_1)

```

```{r}
#Calculate the lm of total carbon per tree in relation to their size
names(mangrove_2017)

lm_carbon <- lm(total_tree_kg_c ~ cd_chatting_m, data=mangrove_2017) |> 
  summary()


```

```{r}

crit_val_95 <- qnorm(0.025, lower.tail = FALSE)
PE_carbon <- lm_carbon$coefficients[,"Estimate"][2]
SE_carbon <- lm_carbon$coefficients[,"Std. Error"][2]

ci_lower = round(PE_carbon - crit_val_95 * SE_carbon, 2)
ci_upper = round(PE_carbon + crit_val_95 * SE_carbon, 2)

ci_lower
ci_upper

```
