---
title: "EDS 222: Final Project"
subtitle: "Mangrove Blue Carbon"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(pwr)

```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")

```

Reading the .csv files

```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","clean_monitoring_data_javier.csv")) |> clean_names()

mangrove_df_b <- read_csv(here::here("data","monitoring_siteB_data_javier.csv")) |> clean_names()

mangrove_df_b <- read_csv(here::here("data","monitoring_siteB_data_javier.csv")) |> clean_names()



```

Understand the data with R methods
```{r}
#Wranggle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)

```


#Cleaning and adding some changes to the data
```{r}
#Changing some data
mangrove_df$height_cm[1253] <- 112
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)

```

Create a table that is grouped by the each plantation year and contains the total number of sample plots and the total number of samples trees per plantation.

```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n())
          

knitr::kable(mangrove_summary)

```

Create a table that shows the samples per each plot

```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot) |> 
  summarise(plot = unique(plot),
            sample_count = n())  

knitr::kable(plots)


```

Create a table of samples per plot but grouping by sample size

```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_m = mean(cd_chatting_m),
            mean_tot_carbon_kg_c = mean(total_tree_kg_c))

kable(mangrove_plots_summary)

```

Create plots that are relevant to your questions

Carbon Sequastration (kg/tree) vs Size
(_ cd_chatting_m, y = total_tree_kg_c _)
```{r}
#Plotting variables of interest
ggplot(mangrove_df, aes(x=cd_chatting_m, y = total_tree_kg_c, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Carbon Sequastration (kg/tree)  vs Size")
```



Lets focus on the 2017 plantation year

Why?
- We have more data
- There are 3 different sampling collection efforts.

```{r}
# Potential need to rename the plots as "Sample Plot#1", "Sample plot #2", etc
# Potential need to rename the plot size as "25m2", "77m2", "154m2"

mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  filter(total_tree_mg_c_ha < 1.9) |> 
  select(height_cm,plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha)

kable(mangrove_2017)


```

Create a table that contains the means of carbon and carbon per hectare for each type plot and each plot
```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n(),
            mean_carbon = mean(total_tree_kg_c),
            mean_carbon_ha = mean(total_tree_mg_c_ha))

kable(plots_2017)

```

```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")
```

-   Create a table and calculate the mean and sd on total carbon per hectare for each plot size

```{r}
#Statistical Table with interesting variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_chatting_size_m = mean(cd_chatting_m),
            mean_tree_c_kg = mean(total_tree_kg_c),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_chatting_size_m = sd(cd_chatting_m),
            sd_tree_c_kg = mean(total_tree_kg_c),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

kable(mangrove_plots_2017)
```

Violin Chart
Tree Heights effects on carbon sequestration
```{r}

ggplot(mangrove_2017, aes(x = plot_size_m2, y= height_cm, size= total_tree_kg_c , fill = plot_size_m2)) +
  geom_violin(alpha = 0.5) +
  labs(title= "Tree heights relationship with carbon sequestration per Plot Type")

```

Violin Chart

Sample size effects on carbon sequestration per plot size

```{r}

ggplot(plots_2017, aes(x = plot_size_m2, y= sample_count, size= mean_carbon , fill = plot_size_m2)) +
  geom_violin(alpha = 0.5) +
  labs(title= "Sample counts effect on carbon sequestration per Plot Type")

```


Bar Charts
```{r}
positions <- c("25m2", "77m2", "154m2")

ggplot(mangrove_plots_2017, aes(x=plot_size_m2 , y=mean_chatting_size_m, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Tree Size means") + 
  scale_fill_brewer(palette = "Spectral") + scale_x_discrete(limits = positions)
  

ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c_kg, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Carbon mean per tree")+ 
  scale_fill_brewer(palette = "Spectral") + scale_x_discrete(limits = positions)


ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c_ha_mg, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Carbon mean per tree per ha") + 
  scale_fill_brewer(palette = "Spectral") +
  scale_x_discrete(limits = positions)

```


```{r}
ggplot(mangrove_plots_2017, aes(x=plot_size_m2, fill = plot_size_m2)) +
  geom_col(aes(y=mean_chatting_size_m)) + 
  geom_col(aes(y=mean_tree_c_ha_mg)) +
  labs(title = "Tree Size means") + 
  scale_fill_brewer(palette = "Spectral") +
  scale_y_continuous(
    name = "Size",
    sec.axis = sec_axis(~ . * 10, name = "CPH"))


```


Option 1 For variable Selection. 
Total Carbon per tree for Hectare vs Sample Plot Type (Plot Type 25m2 vs 154m2)

Histogram 1 : Plot Size 25m2
```{r}

mangrove_2017 |> 
  filter(plot_size_m2 == 25.133 ) |> 
ggplot() +
  geom_histogram(aes(x=total_tree_mg_c_ha)) 
```

Histogram 2 : Plot Size 154m2
```{r}
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938 ) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram() 
```


```{r}
#Variable Total Carbon per tree (mg)
mangrove_2017 |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha ,color = plot_size_m2)) +
  geom_point(aes(y= mean(total_tree_mg_c_ha)), size = 3, alpha = 0.9, color = "darkblue", shape = 3 ) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2")
```

State your null and alternative hypotheses

The Null Hypothesis - The Total Carbon per Tree (mg) mean in the Sample Size of 25m2 is no different to the 154m2.

Alternative - The Total Carbon per Tree (mg) in the sample size of 25m2 is different than the one to the 154m2.

In other words:
For the null hypothesis you state that there is no difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.

Alternative hypothesis you state that there is a difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.


$$H_0: \mu_{154m2} - \mu_{25m2} = 0$$
Calculate the Point Estimate
```{r}
#Calculate Point Estimate
names(mangrove_plots_2017)
point_est_1 = (mangrove_plots_2017$mean_tree_c_ha_mg[1] - mangrove_plots_2017$mean_tree_c_ha_mg[3])

print(point_est_1)

```

Calculate the Standard Error
```{r}
#Define the Standard Error
n_25m = mangrove_plots_2017[[1,2]]
n_154m = mangrove_plots_2017[[3,2]]

s_25m =  mangrove_plots_2017[[1,9]]
s_154m =  mangrove_plots_2017[[3,9]]

SE_1 = as.numeric(sqrt(s_25m^2/n_25m + s_154m^2/n_154m))
print(SE_1)

```

Calculate the Z-Score
```{r}
z_score_1 <- (point_est_1 - 0) / SE_1
z_score_1
```

Calculate the P-Value
```{r}

option1_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2==25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option1_ttest

```

Check differences between manual and t.test automatic

```{r}

#P-value with T.Test
print(option1_ttest$p.value)

#P-Value with pnorm
pval_1 = 2 * pnorm(point_est_1, mean = 0, sd = SE_1, lower.tail = FALSE)
print(pval_1)
```
**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 25m2 to the sample size of 154m2**

```{r}

crit_val_1 = qnorm(0.025, lower.tail = F)
ci_lower_1 = round(point_est_1 - crit_val_1*SE_1, 2)
ci_upper_1 = round(point_est_1 + crit_val_1*SE_1, 2)


print(paste0("There is 95% probability that between [", ci_lower_1,",",ci_upper_1, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 154m2"))
```


We are 95% confident that the true value lies between 0.31 - 0.37 Tree Carbon sequestration in mg per hectare.


```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(25.133, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(aes(y= mean(total_tree_mg_c_ha)), size = 3, alpha = 0.9, color = "darkblue", shape = 3 ) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0.31, linetype = "dashed", color = "deepskyblue4", size = .5)+
  geom_hline(yintercept = 0.37, linetype = "dashed", color = "orange", size = .5)
```


Option 2 For variable Selection
Total Carbon per tree vs Tree Size
```{r}
#Variable Total Carbon per tree (kg)
mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_line(alpha = 0.3, color = "deepskyblue4") +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size",
       x = "Size (chatt m)",
       y = "Tree Carbon (kg)")
```

```{r}

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c, col = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.2) +
  labs(title = "Tree kg C vs Size") +
  facet_wrap(~ plot_size_m2) +
  scale_color_brewer(palette = "Dark2")

```

Starting the Statistical analysis

```{r}

#Calculate the lm of total carbon per tree in relation to their size

lm_carbon <- lm(cd_chatting_m ~ total_tree_kg_c, data=mangrove_2017) |>
  summary()

crit_val_95 <- qnorm(0.025, lower.tail = FALSE)
PE_carbon <- lm_carbon$coefficients[,"Estimate"][2]
SE_carbon <- lm_carbon$coefficients[,"Std. Error"][2]

ci_lower = round(PE_carbon - crit_val_95 * SE_carbon, 2)
ci_upper = round(PE_carbon + crit_val_95 * SE_carbon, 2)

ci_lower
ci_upper

```


```{r}

t.test(mangrove_2017$cd_chatting_m, mangrove_2017$total_tree_kg_c)

```

**Confident Intervals: We are 95% confident that our point estimate lays within the confidence interval 2.5 - 2.6 of  tree size.**


```{r}

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_line(alpha = 0.3, color = "deepskyblue4") +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size",
       x = "Size (chatt m)",
       y = "Tree Carbon (kg)") +
    geom_hline(yintercept = 2.7, color = "orange")+
  geom_hline(yintercept = 2.5, color = "deepskyblue4")
```


Now same thing but with the logs
```{r}
#Take the log of the columns to make it a linear regression
mangrove_2017_option2 <- mangrove_2017 |> 
  mutate(total_tree_kg_c = log10(total_tree_kg_c),
         cd_chatting_m = log10(cd_chatting_m))

#Plot the linear regresion
mangrove_2017_option2 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_line(alpha = 0.3, color = "deepskyblue4") +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size")

```


```{r}
#Take the lm of the new linear model
mangrove_2017_option2 <- mangrove_2017_option2[is.finite(mangrove_2017_option2$total_tree_kg_c), ]


lm_carbon_2 <- lm(total_tree_kg_c ~ cd_chatting_m, data=mangrove_2017_option2) |>
  summary()

lm_carbon_2

```


Calculate Point Estimate, Standard Error, Point Estimate, Confident Intervals
```{r}

crit_val_95 <- qnorm(0.025, lower.tail = FALSE)
PE_carbon_2 <- lm_carbon_2$coefficients[,"Estimate"][2]
SE_carbon_2 <- lm_carbon_2$coefficients[,"Std. Error"][2]

ci_lower_2 = round(PE_carbon_2 - crit_val_95 * SE_carbon_2, 2)
ci_upper_2 = round(PE_carbon_2 + crit_val_95 * SE_carbon_2, 2)

ci_lower_2
ci_upper_2

```

```{r}

mangrove_2017_option2 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_line(alpha = 0.3, color = "deepskyblue4") +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree kg C vs Size") +
  geom_hline(yintercept = -0.19, color = "orange")+
  geom_hline(yintercept = -0.22, color = "deepskyblue4")
```


```{r}

t.test(mangrove_2017_option2$cd_chatting_m, mangrove_2017_option2$total_tree_kg_c)

ci_lower_2
ci_upper_2

```


Option 3 - Power Calculation
Define the Sample Number with Total Carbon per tree

```{r}
#Using our table mangrove_plots_2017 with the means, sd per sample type we can calculate our variables to define the power needed.

```

Plot the histogram distribution
```{r}
#Total Tree Carbon (mg/ha)
mangrove_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=total_tree_mg_c_ha, fill = plot_size_m2)) +
  geom_histogram() +
  labs(title = "Total Carbon per Tree (mg/ha)")
  
```


```{r}
#Total Tree Carbon (mg/ha)
# Just the 25m2
mangrove_2017 |> 
  filter(plot_size_m2 == 25.133) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "red", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")
```


```{r}
#Total Tree Carbon (mg/ha)
# Just the 77m2

mangrove_2017 |> 
  filter(plot_size_m2 == 76.969) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "green", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")
```


```{r}

#Total Tree Carbon (mg/ha)
# Just the 154m2
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightblue", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")

```

Plot the histogram removing the points far away
```{r}
mangrove_2017 |>
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_boxplot() +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") 
```

Create a random table of 10 samples per plot size
```{r}
#Create the random table

random_table <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = 10)

ggplot(random_table, aes(x = total_tree_mg_c_ha, fill = plot_size_m2)) + 
  geom_histogram()

```


Calculate power
```{r}
#Create a table with means and SD per plot size
power_table <- random_table |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_chatting_size_m = mean(cd_chatting_m),
            mean_tree_c_kg = mean(total_tree_kg_c),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_chatting_size_m = sd(cd_chatting_m),
            sd_tree_c_kg = mean(total_tree_kg_c),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))
```


```{r}
# Calculate power with the random samples of 10

random_sd_25m = power_table[[1,9]]
random_sd_154m =  power_table[[3,9]]

alpha = 0.05
power = 0.90
random_d = as.numeric(sqrt(random_sd_25m^2/10 + random_sd_154m^2/10))
print(random_d)

pwr.t.test(n =10, d = random_d, sig.level=.05, alternative = "greater")

```


```{r}

# Calculate the power with all data
sd_25m = mangrove_plots_2017[[1,9]]
sd_154m =  mangrove_plots_2017[[3,9]]

alpha = 0.05
power = 0.90
d = as.numeric(sqrt(sd_25m^2/2 + sd_154m^2/2))
print(d)

#Use pwr.t2n.test when the sample size are not the same
pwr.t2n.test(n1 = 321, n2= 3467, d=d )

```

