---
title: "EDS 222: Final Project"
subtitle: "Mangrove Blue Carbon"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(pwr)

```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")

```

Reading the .csv files

```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","clean_monitoring_data_javier.csv")) |> clean_names()

mangrove_df_b <- read_csv(here::here("data","monitoring_siteB_data_javier.csv")) |> clean_names()

```

Understand the data with R methods
```{r}
#Wranggle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)

```


#Cleaning and adding some changes to the data
```{r}
#Changing some data
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)


```

Create a table that is grouped by the each plantation year containing the total number of sample plots and the total number of samples trees per plantation.

```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n())
          

knitr::kable(mangrove_summary)

```

Create a table that shows the samples per each plot

```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot,plot_size_m2) |> 
  summarise(plot = unique(plot),
            sample_count = n())

```

```{r}

ggplot(plots, aes(x= sample_count, y =plot_size_m2, color = plantation_year)) +
  geom_point() 

  
```


```{r}

ggplot(plots_2017, aes(x = sample_count, y =mean_carbon_mg_ha, color =plot_size_m2 )) +
  geom_point()

```


Create a table of samples per plot but grouping by sample size

```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_cm = (mean(cd_chatting_m)),
            mean_tot_carbon_g_c = (mean(total_tree_kg_c)),
            sd_plant_size = sd(cd_chatting_m),
            sd_tot_carbon_kg_c = sd(total_tree_kg_c))

kable(mangrove_plots_summary)

```

Create plots that are relevant to your questions

Carbon Sequestration (kg/tree) vs Size

(_ cd_chatting_m, y = total_tree_kg_c _)
```{r}
#Plotting variables of interest
ggplot(mangrove_df, aes(x=cd_chatting_m, y = total_tree_kg_c, color = plantation_year)) +
  facet_wrap(~plantation_year) +
  geom_point(alpha = 0.7) +
  labs(title = "Carbon Sequastration (kg/tree)  vs Size")
```



Lets focus on the 2017 plantation year

Why?
- We have more data
- There are 3 different sampling collection efforts.

```{r}
# Potential need to rename the plots as "Sample Plot#1", "Sample plot #2", etc
# Potential need to rename the plot size as "25m2", "77m2", "154m2"

mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  filter(total_tree_mg_c_ha < 1.9) |> 
  select(height_cm,plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha)

```

Create a table that contains the means of carbon and carbon per hectare for each type plot and each plot
```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n(),
            mean_tree_c = mean(total_tree_kg_c),
            mean_carbon_mg_ha = mean(total_tree_mg_c_ha),
            sd_tree_c = sd(total_tree_kg_c),
            sd_tree_c_ha = sd(total_tree_mg_c_ha))

```

```{r}

library(ggpmisc)
my.formula <- y ~ x

ggplot(plots_2017, aes(x= sample_count, y = sd_tree_c_ha)) +
  geom_point(aes(color = plot_size_m2)) +
  geom_smooth(method = lm,
              formula = my.formula,
              color = "cyan4",
              se = F) +
  labs(title = "Tree Carbon variance per Sample plot",
       x = "Samples per Plot",
       y = "Variance of Tree carbon (mg/ha)")


```

```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")

```

-   Create a table and calculate the mean and sd on total carbon per hectare for each plot size

```{r}
#Statistical Table with interesting variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_size_m = mean(cd_chatting_m),
            mean_tree_c = mean(total_tree_kg_c),
            mean_tree_c_ha = mean(total_tree_mg_c_ha),
            sd__size_m = sd(cd_chatting_m),
            sd_tree_c = mean(total_tree_kg_c),
            sd_tree_c_ha = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

kable(mangrove_plots_2017)

```

Violin Chart
Tree Heights effects on carbon sequestration
```{r}

ggplot(mangrove_2017, aes(x = plot_size_m2, y= height_cm, size= total_tree_kg_c , fill = plot_size_m2)) +
  geom_violin(alpha = 0.5) +
  labs(title= "Tree heights relationship with carbon sequestration per Plot Type")

```

Violin Chart
Tree size effects on carbon sequestration

```{r}
ggplot(mangrove_2017, aes(x = plot_size_m2, y= cd_chatting_m, size= total_tree_kg_c , fill = plot_size_m2)) +
  geom_violin(alpha = 0.5) +
  labs(title= "Tree heights relationship with carbon sequestration per Plot Type")
```


Violin Chart
Sample size effects on carbon sequestration per plot size

```{r}

ggplot(plots_2017, aes(x = plot_size_m2, y= sample_count, size= mean_carbon_mg_ha , fill = plot_size_m2)) +
  geom_violin(alpha = 0.5) +
  labs(title= "Sample counts effect on carbon sequestration per Plot Type")

```


Bar Charts
```{r}
positions <- c("25m2", "77m2", "154m2")

ggplot(mangrove_plots_2017, aes(x=plot_size_m2 , y=mean_size_m, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Tree Size means") + 
  scale_fill_brewer(palette = "Set2") + scale_x_discrete(limits = positions)
  

ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Carbon mean per tree")+ 
  scale_fill_brewer(palette = "Spectral") + scale_x_discrete(limits = positions)


ggplot(mangrove_plots_2017, aes(x=plot_size_m2, y=mean_tree_c_ha, fill = plot_size_m2)) +
  geom_col() +
  labs(title = "Carbon mean per tree per ha") + 
  scale_fill_brewer(palette = "Spectral") +
  scale_x_discrete(limits = positions)

```


Compare the 3 histograms of plot size m2
```{r}

ggplot(mangrove_2017, aes(x=total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_freqpoly(aes(y = stat(density))) + 
  scale_color_brewer(palette = "Set1")

```

Option 1 For variable Selection. 
Total Carbon per tree for Hectare vs Sample Plot Type (Plot Type 25m2 vs 154m2)

Histogram 1 : Plot Size 25m2
```{r}

mangrove_2017 |> 
  filter(plot_size_m2 == 25.133 ) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), 
fill = "lightgreen",
color = "gray30", 
alpha = 0.2) +
  geom_density()


```

Histogram 2 : Plot Size 154m2
```{r}

mangrove_2017 |> 
  filter(plot_size_m2 == 153.938 ) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), 
fill = "lightblue",
color = "gray30", 
alpha = 0.2) +
  geom_density()
```

Option 1:
State your null and alternative hypotheses

The Null Hypothesis - The Total Carbon per Tree (mg) mean in the Sample Size of 25m2 is no different to the 154m2.

Alternative - The Total Carbon per Tree (mg) in the sample size of 25m2 is different than the one to the 154m2.

In other words:
For the null hypothesis you state that there is no difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.

Alternative hypothesis you state that there is a difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.


$$H_0: \mu_{25m2} - \mu_{154m2} = 0$$
Calculate the Point Estimate
```{r}
#Calculate Point Estimate

point_est_1 = (mangrove_plots_2017$mean_tree_c_ha_mg[1] - mangrove_plots_2017$mean_tree_c_ha_mg[3])

print(point_est_1)

```

Calculate the Standard Error
```{r}
#Define the Standard Error
n_25m = mangrove_plots_2017[[1,2]]
n_154m = mangrove_plots_2017[[3,2]]

s_25m =  mangrove_plots_2017[[1,9]]
s_154m =  mangrove_plots_2017[[3,9]]

SE_1 = as.numeric(sqrt(s_25m^2/n_25m + s_154m^2/n_154m))
print(SE_1)

```

Calculate the Z-Score
```{r}
z_score_1 <- (point_est_1 - 0) / SE_1
z_score_1

```

Calculate the P-Value
```{r}

option1_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2==25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option1_ttest

```

Check differences between manual and t.test automatic

```{r}

#P-value with T.Test
print(option1_ttest$p.value)

#P-Value with pnorm
pval_1 = pnorm(point_est_1, mean = 0, sd = SE_1, lower.tail = FALSE)

print(pval_1)

```
**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 25m2 to the sample size of 154m2**

```{r}

crit_val_1 = qnorm(0.025, lower.tail = F)
ci_lower_1 = round(point_est_1 - crit_val_1*SE_1, 2)
ci_upper_1 = round(point_est_1 + crit_val_1*SE_1, 2)


print(paste0("There is 95% probability that between [", ci_lower_1,",",ci_upper_1, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 154m2"))

```


We are 95% confident that the true value lies between 0.30 - 0.36 Tree Carbon sequestration in mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(25.133, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(alpha = 0.5) +
  stat_summary(fun= "mean", aes(shape= "mean"),color = "darkblue", geom = "pointrange",size = 1.5) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") +
  geom_hline(aes(yintercept = 0.31, linetype = "Lower CI"), color = "gray50", size = .5) +
  geom_hline(aes(yintercept = 0.37, linetype = "Upper CI"), color = "gray10", size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Confidence Intervals", values = c(2, 2), 
                        guide = guide_legend(override.aes = list(color = c("gray50", "gray10"))))

```


**77m2 vs 154m2** **The Null Hypothesis -** The Total Carbon per Tree (mg/ha) mean in the Sample Size of 77m2 is no different to the 154m2.

**Alternative Hypothesis-** The Total Carbon per Tree (mg/ha) in the sample size of 77m2 is different than the one to the 154m2.

$$H_0: \mu_{77m2} - \mu_{154m2} = 0$$

Calculate the Point Estimate

```{r}
#Calculate Point Estimate

point_est_2 = (mangrove_plots_2017$mean_tree_c_ha[2] - mangrove_plots_2017$mean_tree_c_ha[3])


print(point_est_2)

```

Calculate the Standard Error

```{r}
#Define the Standard Error
n_77m = mangrove_plots_2017[[2,2]]

s_77m =  mangrove_plots_2017[[2,8]]

SE_2 = as.numeric(sqrt(s_77m^2/n_77m + s_154m^2/n_154m))
print(SE_2)

```

Calculate the Z-Score

```{r}
z_score_2 <- (point_est_2 - 0) / SE_2
z_score_2

```

Calculate the P-Value

```{r}

option2_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 76.969], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option2_ttest

```

Check differences between manual and t.test automatic

```{r}

#P-value with T.Test
print(option2_ttest$p.value)

#P-Value with pnorm
pval_2 = pnorm(point_est_2, mean = 0, sd = SE_2, lower.tail = FALSE)

print(pval_2)
```

**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 77m2 to the sample size of 154m2**

```{r}

crit_val_2 = qnorm(0.025, lower.tail = F)
ci_lower_2 = round(point_est_2 - crit_val_2*SE_2, 2)
ci_upper_2 = round(point_est_2 + crit_val_2*SE_2, 2)


print(paste0("There is 95% probability that between [", ci_lower_2,",",ci_upper_2, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 77m2 and 154m2"))

```

We are 95% confident that the true value lies between 0.03 - 0.04 Tree Carbon sequestration in mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(76.969, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(alpha = 0.5) +
  stat_summary(fun= "mean", aes(shape= "mean"),color = "darkblue", geom = "pointrange",size = 1.5) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") +
  geom_hline(aes(yintercept = 0.03, linetype = "Lower CI"), color = "gray50", size = .5) +
  geom_hline(aes(yintercept = 0.04, linetype = "Upper CI"), color = "gray10", size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Confidence Intervals", values = c(2, 2), 
                        guide = guide_legend(override.aes = list(color = c("gray50", "gray10"))))

```

**25m2 vs 77m2** **The Null Hypothesis -** The Total Carbon per Tree (mg/ha) mean in the Sample Size of 25m2 is no different to the 77m2.

**Alternative Hypothesis-** The Total Carbon per Tree (mg/ha) in the sample size of 25m2 is different than the one to the 77m2.

$$H_0: \mu_{77m2} - \mu_{154m2} = 0$$

Calculate the Point Estimate

```{r}
#Calculate Point Estimate

point_est_3 = (mangrove_plots_2017$mean_tree_c_ha[1] - mangrove_plots_2017$mean_tree_c_ha[2])

print(point_est_3)

```

Calculate the Standard Error

```{r}


SE_3 = as.numeric(sqrt(s_25m^2/n_25m + s_77m^2/n_77m))
print(SE_3)

```

Calculate the Z-Score

```{r}
z_score_3 <- (point_est_3 - 0) / SE_3
z_score_3

```

Calculate the P-Value

```{r}

option3_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 76.969])

option3_ttest

```

Check differences between manual and t.test automatic

```{r}

#P-value with T.Test
print(option3_ttest$p.value)

#P-Value with pnorm
pval_3 = pnorm(point_est_3, mean = 0, sd = SE_3, lower.tail = FALSE)

print(pval_3)
```

**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 77m2 to the sample size of 154m2**

```{r}

crit_val_3 = qnorm(0.025, lower.tail = F)
ci_lower_3 = round(point_est_3 - crit_val_3*SE_3, 2)
ci_upper_3 = round(point_est_3 + crit_val_3*SE_3, 2)


print(paste0("There is 95% probability that between [", ci_lower_3,",",ci_upper_3, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 77m2"))

```

We are 95% confident that the true value lies between 0.28 - 0.34 Tree Carbon sequestration in mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(25.133, 76.969)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(aes(y= mean(total_tree_mg_c_ha)), size = 3, alpha = 0.9, color = "darkblue", shape = 3 ) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type")+
  scale_color_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0.28, linetype = "dashed", color = "deepskyblue4", size = .5)+
  
  geom_hline(yintercept = 0.34, linetype = "dashed", color = "orange", size = .5)

```





##Option 2

$$tree carbon_i=\beta_{0}+\beta_{1} \cdot tree size_i +\varepsilon_i$$
Create a graph of two numerical axis to create an analysis

```{r}
# Tree Size vs. Tree Carbon (No Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree Carbon (kg) vs Size",
       x = "Size (chatt m)",
       y = "Tree Carbon (kg)")

```


Using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ using this sample of data.

```{r}

lm(total_tree_kg_c ~ cd_chatting_m, data=mangrove_2017) %>%
  summary()


```

Adding the Interaction

$$tree carbon_i=\beta_{0}+\beta_{1} \cdot tree size_i +\beta_{2} \cdot \text plot size +\varepsilon_i$$

```{r}
# Tree Size vs. Tree Carbon (Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree Carbon (mg/ha) vs Size",
       x = "Tree Size (m)",
       y = "Tree Carbon (mg/ha)") +
  scale_color_brewer(palette = "Dark2")

```

using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}
lm(total_tree_mg_c_ha ~ cd_chatting_m + plot_size_m2, data=mangrove_2017) %>%
  summary() 


```


Option 2 With log

```{r}
#Take the log of the columns to make it a linear regression
mangrove_2017_log <- mangrove_2017 |> 
  mutate(total_tree_kg_c = log10(total_tree_kg_c),
    total_tree_mg_c_ha = log10(total_tree_mg_c_ha),
         cd_chatting_m = log10(cd_chatting_m))

```


```{r}
# No Interaction plot
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Log Tree Carbon (kg) vs Size") 

```

using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}

lm_log_kg <- mangrove_2017_log[is.finite(mangrove_2017_log$total_tree_kg_c), ]

lm(total_tree_kg_c ~ cd_chatting_m + plot_size_m2, data=lm_log_kg) |>
  summary()


```

```{r}
#Plot the linear regression
# Interaction
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.8) +
  labs(title = "Tree kg C vs Size") 

```


Using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}

lm_log_ha <- mangrove_2017_log[is.finite(mangrove_2017_log$total_tree_mg_c_ha), ]

lm(total_tree_kg_c ~ cd_chatting_m + plot_size_m2, data=lm_log_ha) |>
  summary()

```


# INVALID AUGMENT
```{r}
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.3) +
  geom_line(data = augment(lm_mangrove), aes(y = .fitted, color = plot_size_m2)) +
  labs(title = "Tree kg C vs Size")

```




Option 3 - Power Calculation
Define the Sample Number with Total Carbon per tree

Using our table mangrove_plots_2017 with the means, sd per sample type we can calculate our variables to define the power needed.

State your null and alternative hypotheses

The Null Hypothesis - The Total Carbon per Tree (mg) mean in the Sample Size of 25m2 is no different to the 154m2.

Alternative - The Total Carbon per Tree (mg) in the sample size of 25m2 is different than the one to the 154m2.

Question: When we have different categorical variables can you state a hypothesis will all of the variables or you need to specify the difference between two?

Plot the histogram distribution
```{r}
#Total Tree Carbon (mg/ha)
mangrove_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=total_tree_mg_c_ha, fill = plot_size_m2)) +
  geom_histogram(alpha = 0.8) +
  labs(title = "Total Carbon per Tree (mg/ha)")

ggplot(mangrove_2017, aes(x=total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_freqpoly(aes(y = stat(density)))
  
```

**The problem that we see in this histograms is that the means overlaps between the plot size 154m2 and the plot size 77m2. In this case we will need more samples in order to get the 95% confidence we are looking for.

```{r}
#Total Tree Carbon (mg/ha)
# Just the 25m2
mangrove_2017 |> 
  filter(plot_size_m2 == 25.133) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightgreen", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")
```


```{r}
#Total Tree Carbon (mg/ha)
# Just the 77m2
mangrove_2017 |> 
  filter(plot_size_m2 == 76.969) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightblue", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")
```


```{r}
#Total Tree Carbon (mg/ha)
# Just the 154m2
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "pink", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)")

```

In this box plot we can see how is the variance "means" merging from one plot size to the other.
```{r}
mangrove_2017 |>
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_boxplot() +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") 


```

Create a random table of 10 samples per plot size.
```{r}
#Create the random table

random_table <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = 10)

rand_means <- random_table |> 
  group_by(plot_size_m2) |> 
  summarise(mean = mean(total_tree_mg_c_ha))

```



Create a histogram too visualize the new 10 random numbers and see the overlap between the two sample sizes. This will help us understand the power calculation. Remember that if the overlap is high then you need a higher number of samples to increase your probability. 

```{r}
random_table <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = 10)

rand_means <- random_table |> 
  group_by(plot_size_m2) |> 
  summarise("mean tree carbon (mg/ha)"= mean(total_tree_mg_c_ha),
            "sd tree carbon (mg/ha)" = sd(total_tree_mg_c_ha))

ggplot(random_table, aes(x = total_tree_mg_c_ha)) + 
  geom_histogram(aes(y = stat(density), 
                 fill = plot_size_m2),
                 alpha = 0.6) +
  geom_density(aes(fill = plot_size_m2, 
               col = plot_size_m2),
               alpha = 0.2) +
  geom_vline(xintercept = rand_means[[1,2]], color = "red") +
    geom_vline(xintercept = rand_means[[2,2]], color = "green") +
    geom_vline(xintercept = rand_means[[3,2]], color = "blue") +
  ylim(0, 30) +
  xlim(0,1.5)

```

DEFINE POWER

Use the pwr package data to define the power
```{r}
# Create the objects with the sd of each group
sd_25m = mangrove_plots_2017[[1,8]]
sd_77m = mangrove_plots_2017[[2,8]]
sd_154m =  mangrove_plots_2017[[3,8]]
mean_25m = mangrove_plots_2017[[1,6]]
mean_77m = mangrove_plots_2017[[2,6]]
mean_154m = mangrove_plots_2017[[3,6]]
```


If you plan to use a two-sample t-test to compare two means, you would use the pwr.t.test function for estimating sample size or power.

A) Power between 25m2 and 154m2
```{r}

mean_difference_a = mean_25m - mean_154m
d25_154 = as.numeric(sqrt(sd_25m^2/2 + sd_154m^2/2))

effect_size_a = mean_difference_a / d25_154

power_test_1 <- pwr.t.test(d = round(effect_size_a,2), 
                           power = 0.90, 
                           sig.level = 0.05,
                           type="two.sample",
                           alternative="two.sided")
print(power_test_1)
plot(power_test_1)
```

Because the means between 25m2 and 154m2 are so different we will only need 10 samples to have a power of 90%

B) Power between 77m2 and 154m2
```{r}
mean_difference_b = mean_77m - mean_154m
d77_154 = as.numeric(sqrt(sd_77m^2/2 + sd_154m^2/2))
effect_size_b = mean_difference_b / d77_154


power_test_2 <- pwr.t.test(d = round(effect_size_b,2), 
                           power = 0.90, 
                           sig.level = 0.05,
                           type="two.sample",
                           alternative="two.sided")
print(power_test_2)
plot(power_test_2)

```
In this case as the means between 77m2 and 154m2 are closer together we can define that with 50 samples we could have a power of 90%.


# Test with Random Samples
Create power table with the means and sd of your random table and run a power analysis with the selected random samples to test my results

Test your results
```{r}

samples = 300
random_table_test <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = samples)

power_table_test <- random_table_test |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

random_mean_25m = power_table_test[[1,4]]
random_mean_77m = power_table_test[[2,4]]
random_mean_154m = power_table_test[[3,4]]
random_sd_25m = power_table_test[[1,5]]
random_sd_77m = power_table_test[[2,5]]
random_sd_154m = power_table_test[[3,5]]

#Using all the samples
#Real Difference 
random_effect_size_a = (random_mean_25m - random_mean_154m) / (as.numeric(sqrt(random_sd_25m^2/2 + random_sd_154m^2/2)))


print(effect_size_a)
print(random_effect_size_a)


```


