---
title: "new-final"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(pwr)
library(xtable)
library(broom)
```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")
```

##Read, clean and understand the data

Reading the .csv file
```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","clean_monitoring_data_javier.csv")) |> clean_names()
```

Understanding the data with different functions
```{r}
#Wrangle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)
str(mangrove_2017)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)

```

## Cleaning and adding some changes to the data
```{r}
#Changing some data
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)

```


## Creating tables 

Create a table that is grouped by each plantation year, showing the total number of sample plots and the total number of trees per plantation effort.
```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise("Plot Count" = length(unique(plot)),
          "Sample Count" = n(),
          "Tree Height mean" = mean(height_cm),
          "Total Carbon (mg/ha)" = sum(total_tree_mg_c_ha))

knitr::kable(mangrove_summary)
```

Create a table that shows the samples per each plot
```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot,plot_size_m2) |> 
  summarise(plot = unique(plot),
            sample_count = n(),
            sd_tree_c_ha = sd(total_tree_mg_c_ha))

```

Create a table of samples per plot but grouping them by sample size
```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_cm = (mean(cd_chatting_m)),
            mean_tot_carbon_g_c = (mean(total_tree_kg_c)),
            sd_plant_size = sd(cd_chatting_m),
            sd_tot_carbon_kg_c = sd(total_tree_kg_c))

kable(mangrove_plots_summary)

```

## The 2017 Decision
Now lets focus on the 2017 plantation year and define our questions

#### Why?
- We have enough relevant data. 
- It has the highest variance from the three replantation years. 
- There are 3 different sampling collection efforts.

Create a table that only contains information from the 2017 plantation year and select the relevant columns.
```{r}
# Potential need to rename the plots as "Sample Plot#1", "Sample plot #2", etc
# Potential need to rename the plot size as "25m2", "77m2", "154m2"

mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  select(height_cm,plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha)

mangrove_2017 <- mangrove_2017[!(mangrove_2017$total_tree_mg_c_ha== 1.224),]

```

Create a table that contains the means of carbon and carbon per hectare for each type plot and each plot
```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n(),
            mean_hight = mean(height_cm),
            mean_tree_c = mean(total_tree_kg_c),
            mean_carbon_mg_ha = mean(total_tree_mg_c_ha),
            sd_tree_c = sd(total_tree_kg_c),
            sd_tree_c_ha = sd(total_tree_mg_c_ha))

```

Create a table that shows percentages of the 2017 replantation effort.
```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")

```

Create a table and calculate the mean and sd on total carbon per hectare for each plot size
```{r}
#Statistical Table with interesting variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_tree_c_ha = mean(total_tree_mg_c_ha),
            sd_tree_c_ha = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

kable(mangrove_plots_2017)

```

Graph the effect of tree size and total carbon

$$tree carbon_{kg}=\beta_{0}+\beta_{1} \cdot tree size_i +\varepsilon_i$$

Create a graph of two numerical axis to create an analysis

```{r}
# Tree Size vs. Tree Carbon (No Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree Size Effect of Total Carbon",
       x = "Tree Size (chatt m)",
       y = "Tree Carbon (kg)")

```

Adding the Interaction

$$tree carbon_{kg/ha} =\beta_{0}+\beta_{1} \cdot tree size_i +\beta_{2} \cdot \text plot size +\varepsilon_i$$

```{r}
# Tree Size vs. Tree Carbon (Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree  Size Effect in Total Carbon per Ha",
       x = "Tree Size (chatt m)",
       y = "Tree Carbon (mg/ha)") +
  scale_color_brewer(palette = "Dark2")

```


##Analyzing the different options to answer our initial research question:

### Questions 1: 

##### Tree Carbon (mg/ha) vs. Sample Types - Create your hypothesis testing, and define the confidence intervals for the Tree Carbon sequestration in mg per hectare depending on the sample type comparison.


Plot Size vs. Total Tree Carbon
```{r}
#Boxplot
mangrove_2017 |>
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_boxplot() +
  labs(title = "Tree Carbon vs Plot Type",
       subititle = "Plantation Year - 2017",
       x = "Plot Size (m2)",
       y = "Total Tree Carbon (mg/ha)") +
  scale_color_brewer(palette = "Dark2") 

```

State your null and alternative hypotheses

**25m2 vs 154m2** 

**The Null Hypothesis-** The Total Carbon per Tree (mg/ha) mean in the Sample Size of 25m2 is no different to the 154m2.

**Alternative Hypothesis-** The Total Carbon per Tree (mg/ha) in the sample size of 25m2 is different than the one to the 154m2.

$$H_0: \mu_{25m2} - \mu_{154m2} = 0$$
$$H_0: \mu_{25m2} - \mu_{154m2} \neq 0$$


Calculate the Point Estimate

```{r}
#Calculate Point Estimate

point_est_1 = (mangrove_plots_2017$mean_tree_c_ha[1] - mangrove_plots_2017$mean_tree_c_ha[3])
print(point_est_1)

```

Calculate the Standard Error
$$SE = \sqrt{\frac{s_1^2}{n_1} + \frac{s^2_2}{n_2}}$$
```{r}
#Define the Standard Error

n_154m = mangrove_plots_2017$tot_samples[3]


s_154m =  mangrove_plots_2017$sd_tree_c_ha[3]

SE_1 = as.numeric(sqrt(s_25m^2/n_25m + s_154m^2/n_154m))
print(SE_1)

```

Calculate the Z-Score
$$z_{score}=\frac{\text { point estimate }-\text { null value }}{S E}$$

```{r}
z_score_1 <- (point_est_1 - 0) / SE_1
z_score_1

```
**Thanks to this calculation we can tell the observed difference in Tree Carbon (mg/ha) is 20.2 standard deviations above our null hypothesis of "zero difference".**

Calculate the T-Test and P-Value
```{r}

option1_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option1_ttest

```

**With this results we can reject the null as the p-value is very small, meaning there is a significant statistical difference between the means on Carbon storage in each tree per hectare between the sample plot size of 25m2 to the sample size of 154m2**


Now lets calculate the Confidence Intervals
```{r}

crit_val_1 = qnorm(0.025, lower.tail = F)
ci_lower_1 = round(point_est_1 - crit_val_1*SE_1, 3)
ci_upper_1 = round(point_est_1 + crit_val_1*SE_1, 3)


print(paste0("There is 95% probability that between [", ci_lower_1,",",ci_upper_1, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 154m2"))

```

We are 95% confident that the true value of the difference in tree carbon across the two plot sizes lies between 0.308 - 0.374 mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(25.133, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(alpha = 0.5) +
  stat_summary(fun= "mean", aes(shape= "mean"),color = "darkblue", geom = "point",size = 5) +
  labs(title = "Tree Carbon (mg/ha) vs Plot Type") +
  scale_color_brewer(palette = "Dark2") +
  geom_hline(aes(yintercept = 0.31, linetype = "Lower CI"), color = "gray50", size = .5) +
  geom_hline(aes(yintercept = 0.37, linetype = "Upper CI"), color = "gray10", size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Confidence Intervals", values = c(2, 2), 
                        guide = guide_legend(override.aes = list(color = c("gray50", "gray10"))))

```


Now, the difference of the means between a sample size plot of 25m2 vs. a sample plot of 154m2 are very different. BUT between a sample plot of 77m2 and the sample plot of 154m2 the mean seems to be way closer. Now, lets analyse this two more similar types of plot.



**77m2 vs 154m2** **The Null Hypothesis -** The Total Carbon per Tree (mg/ha) mean in the Sample Size of 77m2 is no different to the 154m2.

**Alternative Hypothesis-** The Total Carbon per Tree (mg/ha) in the sample size of 77m2 is different than the one to the 154m2.

$$H_0: \mu_{77m2} - \mu_{154m2} = 0$$

$$H_0: \mu_{77m2} - \mu_{154m2} \neq 0$$

Calculate the Point Estimate

```{r}
#Calculate Point Estimate

point_est_2 = (mangrove_plots_2017$mean_tree_c_ha[2] - mangrove_plots_2017$mean_tree_c_ha[3])
print(point_est_2)

```

Calculate the Standard Error

```{r}
#Define the Standard Error
n_77m = mangrove_plots_2017$tot_samples[2]
s_77m =  mangrove_plots_2017$sd_tree_c_ha[2]


SE_2 = as.numeric(sqrt(s_77m^2/n_77m + s_154m^2/n_154m))
print(SE_2)

```

Calculate the Z-Score

```{r}
z_score_2 <- (point_est_2 - 0) / SE_2
z_score_2

```
**Now we can tell that the Z-Score between a sample plot of 77m2 and 154m2 is lower, meaning that the the observed difference in Tree Carbon (mg/ha) is 9.23 standard deviations above our null hypothesis of "zero difference".**


Calculate t - test

```{r}

option2_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 76.969], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option2_ttest

```

**With this results we can also reject the null as the p-value is again very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 77m2 to the sample size of 154m2**


Now lets calculate both confidence intervals
```{r}

crit_val_2 = qnorm(0.025, lower.tail = F)
ci_lower_2 = round(point_est_2 - crit_val_2*SE_2, 3)
ci_upper_2 = round(point_est_2 + crit_val_2*SE_2, 2)


print(paste0("There is 95% probability that between [", ci_lower_2,",",ci_upper_2, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 77m2 and 154m2"))

```

We are 95% confident that the true value of our point estimate lies between 0.026 - 0.04 Tree Carbon sequestration in mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(76.969, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(alpha = 0.5) +
  stat_summary(fun= "mean", aes(shape= "mean"),color = "darkblue", geom = "pointrange",size = 1.5) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2") +
  geom_hline(aes(yintercept = 0.026, linetype = "Lower CI"), color = "gray50", size = .5) +
  geom_hline(aes(yintercept = 0.04, linetype = "Upper CI"), color = "gray10", size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Confidence Intervals", values = c(2, 2), 
                        guide = guide_legend(override.aes = list(color = c("gray50", "gray10"))))

```


Conclusion:

As we can see in the second graph, now both means are much closer together but satistically significantly enough difference to tell they are different. With this data analysis and results we can help the actual client have a better understanding on their plot sampling sizes, concluding that smaller plot sizes somehow are getting a large carbon estimate.

## **Question 2:** 

What is the density difference between sample sizes and what is the effect of variance in the calculated Total Tree Carbon (mg/ha).

Creating a new table and graph to easily visualize what is the density per plot type (Number of trees per m2)
```{r}


#Changing the column to numeric
mangrove_df$plot_size_m2_num <- as.numeric(mangrove_df$plot_size_m2)

density_df <- read_csv(here::here("data","clean_monitoring_data_javier.csv")) |> clean_names() |> 
  filter(plantation_year == 2017) |> 
  select(height_cm,plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha) |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n(),
            mean_hight = mean(height_cm),
            mean_tree_c = mean(total_tree_kg_c),
            mean_carbon_mg_ha = mean(total_tree_mg_c_ha),
            sd_tree_c = sd(total_tree_kg_c),
            sd_tree_c_ha = sd(total_tree_mg_c_ha)) |> 
  mutate(density = sample_count/plot_size_m2)


ggplot(density_df, aes(x= as.factor(plot_size_m2), y= density, fill = as.factor(plot_size_m2))) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Density of trees per plot size",
       x= "Plot Size (m2)",
       y = "Density (Tree/m2)",
       col = "Plot Size (m2)") +
  scale_fill_discrete(name = "Plot Type")

```
As we can see in this graph, plot sizes have higher density, but also higher variability, In contrast bigger plot sizes have lower density, and also lower variability.

But, how this affects our calculations and what are the estimated number of samples to have an estimate of the results

In this graph we are looking at how the Carbon mean per plot doesn´t change much when you increase the samples. We can see that the total carbon is more influenced by the density than the number of samples per plot.
```{r}

ggplot(density_df, aes(x= sample_count, y= mean_carbon_mg_ha, color = as.factor(plot_size_m2) )) +
  geom_point() +
  labs(title= "Carbon mean per sample count",
       subtitle = 2017,
       x = "Sample Count",
       y = "Carbon Mean (mh/ha)",
       col ="Plot Size (m2)" ) +
   geom_smooth(method = lm,
              color = "cyan4",
              se = F) 
```


Plot the three distributions
```{r}
#Total Tree Carbon (mg/ha)
mangrove_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density), fill = plot_size_m2),color = "gray30", alpha = 0.3) +
  geom_density(aes(fill = plot_size_m2), col = "gray30", alpha = 0.8) +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "2017")
  
```

```{r}
#Plotting just the lines
ggplot(mangrove_2017, aes(x=total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_freqpoly(aes(y = stat(density)))
```



Comparing the variation of samples per plot

Question: What technique of mangrove sampling will bring a lower variance in Total Tree Carbon calculation (mg/ha).

```{r}
#Total Tree Carbon (mg/ha)
# Just the 25m2
mangrove_2017 |> 
  filter(plot_size_m2 == 25.133) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "red", color = "gray30", alpha = 0.6) +
  geom_density(col = "gray30", alpha = 0.3, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "25m2")
```

```{r}
#Total Tree Carbon (mg/ha)
# Just the 77m2
mangrove_2017 |> 
  filter(plot_size_m2 == 76.969) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightgreen", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "77m2")
```

```{r}
#Total Tree Carbon (mg/ha)
# Just the 154m2
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightblue", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "154m2")

```
Now, let analyse what is the effect of this difference in densities and what is this variance telling us.


Create a plot showing the relationship between sample size and variation of the calculated tree carbon (mg/ha)
```{r}
ggplot(plots_2017, aes(x= sample_count, y = sd_tree_c_ha)) +
  geom_point(aes(color = plot_size_m2)) +
  geom_smooth(method = lm,
              color = "cyan4",
              se = F) +
  labs(title = "Tree Carbon variance per Sample plot",
       x = "Samples per Plot",
       y = "Variance of Tree carbon (mg/ha)")
```

As you can see in the graph above the sample type with the lowest variance is the plot 154m2 with a mean of 0.03 in Standard Deviation, but it has a huge amount of samples. So, now the following question is: what is the recommended sample size to have a power of 90%.


Show the histogram for all variances across the 2017 plantation.
```{r}
ggplot(plots_2017, aes(x=sd_tree_c_ha)) +
  geom_histogram(aes(y = stat(density)), color = "gray20", fill = "darkolivegreen", alpha = 0.7) +
  geom_density(col = "gray30", alpha = 0.15, fill = "green") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "All Plots",
       x = "Tree Carbon Variance (mg/ha)",
       y = "Unique Plots")
```

But now if you separate per plot type you have something different
```{r}
plots_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=sd_tree_c_ha)) +
  geom_histogram(aes(y = stat(density), fill = plot_size_m2),color = "gray30", alpha = 0.3) +
  geom_density(aes(fill = plot_size_m2), col = "gray30", alpha = 0.8) +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "Per Plot",
       x = "Tree Cabron Variance (mg/ha)",
       y = "Unique Plots")

```




Calculate the power and number of samples.


Do you remember the histogram with the all the calculations of total carbon per tree (mg/ha)? Well, as you can see below the sample plots of 77m2 and 154m2 are way more overlapped than the 25m2. 

```{r}
#Total Tree Carbon (mg/ha)
mangrove_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density), fill = plot_size_m2),color = "gray30", alpha = 0.3) +
  geom_density(aes(fill = plot_size_m2), col = "gray30", alpha = 0.8) +
  labs(title = "Total Carbon per Tree (mg/ha)")
  
```



With that un mind we will create an analysis of the power calculations
In this case we wil use the pwr.t.test function for estimating sample size or power, as our data contains a two-sample t-test to compare the two means.

```{r}
# Create the objects with the sd of each group
sd_25m = mangrove_plots_2017[[1,5]]
sd_77m = mangrove_plots_2017[[2,5]]
sd_154m =  mangrove_plots_2017[[3,5]]
mean_25m = mangrove_plots_2017[[1,4]]
mean_77m = mangrove_plots_2017[[2,4]]
mean_154m = mangrove_plots_2017[[3,4]]

```

Define the number of counts `n` by using the plot size of 25m2 vs a plot size of 154m2.
A) Power between 25m2 and 154m2
```{r}

mean_difference_a = mean_25m - mean_154m
d25_154 = as.numeric(sqrt(sd_25m^2/2 + sd_154m^2/2))

effect_size_a = mean_difference_a / d25_154

power_test_1 <- pwr.t.test(d = round(effect_size_a,2), 
                           power = 0.90, 
                           sig.level = 0.05,
                           type="two.sample",
                           alternative="two.sided")
print(power_test_1)
```
Because the means between 25m2 and 154m2 are so different we will only need 10 samples to have a power of 90%.

B) Power between 77m2 and 154m2
```{r}
mean_difference_b = mean_77m - mean_154m
d77_154 = as.numeric(sqrt(sd_77m^2/2 + sd_154m^2/2))
effect_size_b = mean_difference_b / d77_154


power_test_2 <- pwr.t.test(d = round(effect_size_b,2), 
                           power = 0.90, 
                           sig.level = 0.05,
                           type="two.sample",
                           alternative="two.sided")
print(power_test_2)
plot(power_test_2)

```
In this case as the means between 77m2 and 154m2 are closer together we can define that with 50 samples we could have a power of 90%.




# Test with my results with random samples
Create power table with the means and sd of your random table and run a power analysis with the selected random samples to test my results

Test your results 25m2 vs 154m2
```{r}

samples = 10
random_table_test <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = samples)

power_table_test <- random_table_test |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

random_mean_25m = power_table_test[[1,4]]
random_mean_154m = power_table_test[[3,4]]
random_sd_25m = power_table_test[[1,5]]
random_sd_154m = power_table_test[[3,5]]

#Using all the samples
#Real Difference 
random_effect_size_a = (random_mean_25m - random_mean_154m) / (as.numeric(sqrt(random_sd_25m^2/2 + random_sd_154m^2/2)))

print(effect_size_a)
print(random_effect_size_a)

ggplot(random_table_test, aes(x= plot_size_m2, y = total_tree_mg_c_ha, col = plot_size_m2)) +
  geom_point() +
  geom_hline(aes(yintercept = effect_size_a, 
                 linetype = "real"), 
             color = "gray50", 
             size = .5) +
  geom_hline(aes(yintercept = random_effect_size_a, 
                 linetype = "random"), 
             color = "gray10", 
             size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Means difference", 
                        values = c(2, 2), 
                        guide = guide_legend(
                          override.aes = list(
                            color = c("gray50", "gray10")))) +
  labs(title = "Total Cabron per Tree (mg/ha)",
       subtitle = "Random sample size",
       caption = paste("Difference between the real sample means is:",round(mean_difference_a,2),"(mg/ha)" ,"\nDifference between the randomly relected samples mean is:", round((random_mean_25m - random_mean_154m),2), "(mg/ha)"))
  
  
```

Test your results 77m2 vs 154m2
```{r}
samples = 50
random_table_test <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  slice_sample(n = samples)

power_table_test <- random_table_test |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_tree_c_ha_mg = mean(total_tree_mg_c_ha),
            sd_tree_c_ha_mg = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

random_mean_77m = power_table_test[[2,4]]
random_mean_154m = power_table_test[[3,4]]
random_sd_77m = power_table_test[[2,5]]
random_sd_154m = power_table_test[[3,5]]

#Using all the samples
#Real Difference 
random_effect_size_b = (random_mean_77m - random_mean_154m) / (as.numeric(sqrt(random_sd_77m^2/2 + random_sd_154m^2/2)))

print(effect_size_b)
print(random_effect_size_b)

ggplot(random_table_test, aes(x= plot_size_m2, y = total_tree_mg_c_ha, col = plot_size_m2)) +
  geom_point() +
  geom_hline(aes(yintercept = effect_size_b, 
                 linetype = "real"), 
             color = "gray50", 
             size = .5) +
  geom_hline(aes(yintercept = random_effect_size_b, 
                 linetype = "random"), 
             color = "gray10", 
             size = .5) +
  scale_shape_manual("", values= c("mean" = "+")) +
  scale_linetype_manual(name = "Means Difference", 
                        values = c(2, 2), 
                        guide = guide_legend(
                          override.aes = list(
                            color = c("gray50", "gray10")))) +
  labs(title = "Total Cabron per Tree (mg/ha)",
       subtitle = "Random sample size",
       caption = paste("Difference between the real sample means is:",round(mean_difference_b,2),"(mg/ha)" ,"\nDifference between the randomly relected samples mean is:", round((random_mean_77m - random_mean_154m),2), "(mg/ha)"))

```

