---
title: "new-final"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(pwr)
library(xtable)
library(broom)
```

```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")
```

Reading the .csv file

```{r}
#Load data Mangrove Data
mangrove_df <- read_csv(here::here("data","clean_monitoring_data_javier.csv")) |> clean_names()
```

Understanding the data with different functions

```{r}
#Wrangle the data for table A
dim(mangrove_df)
names(mangrove_df)
skimr::skim(mangrove_df)
summary(mangrove_df)

unique(mangrove_df$plantation_year)
unique(mangrove_df$plot_size_m2)

class(mangrove_df$plantation_year)

```

#Cleaning and adding some changes to the data

```{r}
#Changing some data
mangrove_df$plantation_year <- as.character(mangrove_df$plantation_year)
mangrove_df$plot <- as.character(mangrove_df$plot)
mangrove_df$plot_size_m2 <- as.factor(mangrove_df$plot_size_m2)


```

Create a table that is grouped by each plantation year, showing the total number of sample plots and the total number of trees per plantation effort.

```{r}

mangrove_summary <- mangrove_df |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n(),
          "sum_tot_c_mg/ha" = sum(total_tree_mg_c_ha))

          

knitr::kable(mangrove_summary)

```

Create a table that shows the samples per each plot

```{r}

plots <- mangrove_df |> 
  group_by(plantation_year,plot,plot_size_m2) |> 
  summarise(plot = unique(plot),
            sample_count = n(),
            sd_tree_c_ha = sd(total_tree_mg_c_ha))

```

Create a table of samples per plot but grouping by sample size

```{r}
mangrove_plots_summary <- mangrove_df |> 
  group_by(plantation_year, plot_size_m2) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_cm = (mean(cd_chatting_m)),
            mean_tot_carbon_g_c = (mean(total_tree_kg_c)),
            sd_plant_size = sd(cd_chatting_m),
            sd_tot_carbon_kg_c = sd(total_tree_kg_c))

kable(mangrove_plots_summary)

```

Create a plot to demonstrate the variance on total carbon that each tree has in mg/ha vs. sample size

```{r}
ggplot(plots, aes(x= sample_count, y = sd_tree_c_ha)) +
  geom_point(aes(color = plantation_year)) +
  geom_smooth(method = lm,
              color = "cyan4",
              se = F) +
  labs(title = "Tree Carbon variance per Sample plot",
       x = "Samples per Plot",
       y = "Variance of Tree carbon (mg/ha)")

```

Lets focus on the 2017 plantation year

Why? - We have enough relevant data. - It has the highest variance from the three replantation years. - There are 3 different sampling collection efforts.

Create a table that only contains information from the 2017 plantation year and the relevant columns.

```{r}
# Potential need to rename the plots as "Sample Plot#1", "Sample plot #2", etc
# Potential need to rename the plot size as "25m2", "77m2", "154m2"

mangrove_2017 <- mangrove_df |> 
  filter(plantation_year == 2017) |> 
  select(height_cm,plot, plot_size_m2, cd_chatting_m,total_tree_kg_c,total_tree_mg_c_ha)

```

Create a table that contains the means of carbon and carbon per hectare for each type plot and each plot

```{r}
plots_2017 <- mangrove_2017 |> 
  group_by(plot, plot_size_m2) |> 
  summarise(sample_count = n(),
            mean_hight = mean(height_cm),
            mean_tree_c = mean(total_tree_kg_c),
            mean_carbon_mg_ha = mean(total_tree_mg_c_ha),
            sd_tree_c = sd(total_tree_kg_c),
            sd_tree_c_ha = sd(total_tree_mg_c_ha))

```

-   Create a table that shows percentages of the 2017 replantation effort.

```{r}
plots_percent_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(plot_count = length(unique(plot)),
            tot_samples = n()) |> 
  mutate(plot_percentage = round((plot_count/sum(plot_count)*100),2),
         tot_samples_percentage = round((tot_samples / sum(tot_samples)*100),2))

kable(plots_percent_2017,
      caption = "TABLE 5: 2017 Sample Plots Percentages")

```

-   Create a table and calculate the mean and sd on total carbon per hectare for each plot size

```{r}
#Statistical Table with interesting variables
mangrove_plots_2017 <- mangrove_2017 |> 
  group_by(plot_size_m2) |> 
  summarise(tot_samples = n(),
            tot_plot_count = length(unique(plot)),
            mean_height = mean(height_cm),
            mean_size_m = mean(cd_chatting_m),
            mean_tree_c_ha = mean(total_tree_mg_c_ha),
            sd__size_m = sd(cd_chatting_m),
            sd_tree_c_ha = sd(total_tree_mg_c_ha)) |> 
  mutate(plot_size_m2 = c("25m2", "77m2", "154m2"))

kable(mangrove_plots_2017)

```

Analyzing the different options that we can do.

**Option 1:** Tree Carbon (mg/ha) vs. Sample Types - Create your hypothesis testing, and define the confidence intervals for the Tree Carbon sequestration in mg per hectare depending on the sample type comparison.

**Option 2:** Total Carbon per tree vs Tree Size - Exponential regression, take the log and calculate the two confidence intervals.

**Option 3:** Calculate the Power with the information that we have.

## **Option 1:**


```{r}
mangrove_2017 |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha ,color = plot_size_m2)) + 
  geom_point(aes(y= mean(total_tree_mg_c_ha)), size = 5, alpha = 0.9, color = "darkblue", shape = 3 ) +
  geom_point(alpha = 0.3, size = 3) +
  labs(title = "Tree mg C vs Plot Type") +
  scale_color_brewer(palette = "Dark2")

```

State your null and alternative hypotheses

**25m2 vs 154m2** 

**The Null Hypothesis-** The Total Carbon per Tree (mg/ha) mean in the Sample Size of 25m2 is no different to the 154m2.

**Alternative Hypothesis-** The Total Carbon per Tree (mg/ha) in the sample size of 25m2 is different than the one to the 154m2.

In other words: For the null hypothesis you state that there is no difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.

Alternative hypothesis you state that there is a difference in the means of the total carbon per tree (mg) between the sample size of 25m2 and 154m2.

$$H_0: \mu_{25m2} - \mu_{154m2} = 0$$

Calculate the Point Estimate

```{r}
#Calculate Point Estimate

point_est_1 = (mangrove_plots_2017$mean_tree_c_ha[1] - mangrove_plots_2017$mean_tree_c_ha[3])

print(point_est_1)

```

Calculate the Standard Error

```{r}
#Define the Standard Error
n_25m = mangrove_plots_2017$tot_samples[1]
n_154m = mangrove_plots_2017$tot_samples[3]

s_25m =  mangrove_plots_2017$sd_tree_c_ha[1]
s_154m =  mangrove_plots_2017$sd_tree_c_ha[3]

SE_1 = as.numeric(sqrt(s_25m^2/n_25m + s_154m^2/n_154m))
print(SE_1)

```

Calculate the Z-Score

```{r}
z_score_1 <- (point_est_1 - 0) / SE_1
z_score_1

```

Calculate the T-Test and P-Value

```{r}

option1_ttest <- t.test(mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 25.133], mangrove_2017$total_tree_mg_c_ha[mangrove_2017$plot_size_m2 == 153.938])

option1_ttest

```

Check differences between manual and t.test automatic

```{r}

#P-value with T.Test
print(option1_ttest$p.value)

#P-Value with pnorm
pval_1 = pnorm(point_est_1, mean = 0, sd = SE_1, lower.tail = FALSE)

print(pval_1)
```

**With this results we can reject the null as the p-value is very small, meaning there is a significant difference between the means on Carbon storage in each tree per hectare between the sample plot size of 25m2 to the sample size of 154m2**

```{r}

crit_val_1 = qnorm(0.025, lower.tail = F)
ci_lower_1 = round(point_est_1 - crit_val_1*SE_1, 2)
ci_upper_1 = round(point_est_1 + crit_val_1*SE_1, 2)


print(paste0("There is 95% probability that between [", ci_lower_1,",",ci_upper_1, "] it contains the difference in total tree carbon (mg) per hecatre across sample plot 25m2 and 154m2"))

```

We are 95% confident that the true value lies between 0.30 - 0.36 Tree Carbon sequestration in mg per hectare.

```{r}
mangrove_2017 |>
  filter(plot_size_m2 %in% c(25.133, 153.938)) |> 
  ggplot(aes(x=plot_size_m2, y = total_tree_mg_c_ha,color = plot_size_m2)) +
  geom_point(aes(y= mean(total_tree_mg_c_ha)), size = 3, alpha = 0.9, color = "darkblue", shape = 3 ) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree mg C vs Plot Type")+
  scale_color_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0.31, linetype = "dashed", color = "deepskyblue4", size = .5)+
  
  geom_hline(yintercept = 0.37, linetype = "dashed", color = "orange", size = .5)

```

##Option 2

$$tree carbon_i=\beta_{0}+\beta_{1} \cdot tree size_i +\varepsilon_i$$
Create a graph of two numerical axis to create an analysis

```{r}
# Tree Size vs. Tree Carbon (No Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Tree Carbon (kg) vs Size",
       x = "Size (chatt m)",
       y = "Tree Carbon (kg)")

```
Using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ using this sample of data.

```{r}

lm(total_tree_kg_c ~ cd_chatting_m, data=mangrove_2017) %>%
  summary()


```

Adding the Interaction

$$tree carbon_i=\beta_{0}+\beta_{1} \cdot tree size_i +\beta_{2} \cdot \text plot size +\varepsilon_i$$

```{r}
# Tree Size vs. Tree Carbon (Interaction)

mangrove_2017 |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.7) +
  labs(title = "Tree Carbon (mg/ha) vs Size",
       x = "Tree Size (m)",
       y = "Tree Carbon (mg/ha)") +
  scale_color_brewer(palette = "Dark2")

```

using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}
lm(total_tree_mg_c_ha ~ cd_chatting_m + plot_size_m2, data=mangrove_2017) %>%
  summary() 


```


Option 2 With log

```{r}
#Take the log of the columns to make it a linear regression
mangrove_2017_log <- mangrove_2017 |> 
  mutate(total_tree_kg_c = log10(total_tree_kg_c),
    total_tree_mg_c_ha = log10(total_tree_mg_c_ha),
         cd_chatting_m = log10(cd_chatting_m))

```


```{r}
# No Interaction plot
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_kg_c)) +
  geom_point(alpha = 0.8, color = "darkolivegreen") +
  labs(title = "Log Tree Carbon (kg) vs Size") 

```

using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}

lm_log_kg <- mangrove_2017_log[is.finite(mangrove_2017_log$total_tree_kg_c), ]

lm(total_tree_kg_c ~ cd_chatting_m + plot_size_m2, data=lm_log_kg) |>
  summary()


```

```{r}
#Plot the linear regression
# Interaction
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.8) +
  labs(title = "Tree kg C vs Size") 

```


Using `lm()` to estimate $\hat\beta_0$, $\hat\beta_1$ and $\hat\beta_2$ using this sample of data.

```{r}

lm_log_ha <- mangrove_2017_log[is.finite(mangrove_2017_log$total_tree_mg_c_ha), ]

lm(total_tree_kg_c ~ cd_chatting_m + plot_size_m2, data=lm_log_ha) |>
  summary()

```


# INVALID AUGMENT
```{r}
mangrove_2017_log |> 
  ggplot(aes(x=cd_chatting_m, y = total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_point(alpha = 0.3) +
  geom_line(data = augment(lm_mangrove), aes(y = .fitted, color = plot_size_m2)) +
  labs(title = "Tree kg C vs Size")

```


Option 3
Plot the three distributions

```{r}
#Total Tree Carbon (mg/ha)
mangrove_2017 |> 
  filter(plot_size_m2 %in% c(25.133, 76.969, 153.938)) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density), fill = plot_size_m2),color = "gray30", alpha = 0.3) +
  geom_density(aes(fill = plot_size_m2), col = "gray30", alpha = 0.8) +
  labs(title = "Total Carbon per Tree (mg/ha)")
  
```

```{r}
#Plotting just the lines
ggplot(mangrove_2017, aes(x=total_tree_mg_c_ha, color = plot_size_m2)) +
  geom_freqpoly(aes(y = stat(density)))
```

# Option 3 comparing the variation of samples per plot
Question: What technique of mangrove sampling will bring a lower variance in Total Tree Carbon calculation (mg/ha).

```{r}
#Total Tree Carbon (mg/ha)
# Just the 25m2
mangrove_2017 |> 
  filter(plot_size_m2 == 25.133) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "red", color = "gray30", alpha = 0.6) +
  geom_density(col = "gray30", alpha = 0.3, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "25m2")
```

```{r}
#Total Tree Carbon (mg/ha)
# Just the 77m2
mangrove_2017 |> 
  filter(plot_size_m2 == 76.969) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightgreen", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "77m2")
```

```{r}
#Total Tree Carbon (mg/ha)
# Just the 154m2
mangrove_2017 |> 
  filter(plot_size_m2 == 153.938) |> 
ggplot(aes(x=total_tree_mg_c_ha)) +
  geom_histogram(aes(y = stat(density)), fill = "lightblue", color = "gray30") +
  geom_density(col = "gray30", alpha = 0.5, fill = "gray10") +
  labs(title = "Total Carbon per Tree (mg/ha)",
       subtitle = "154m2")

```

The final plot showing the relationship between sample size and variation of the calculated tree carbon (mg/ha)

```{r}
ggplot(plots_2017, aes(x= sample_count, y = sd_tree_c_ha)) +
  geom_point(aes(color = plot_size_m2)) +
  geom_smooth(method = lm,
              color = "cyan4",
              se = F) +
  labs(title = "Tree Carbon variance per Sample plot",
       x = "Samples per Plot",
       y = "Variance of Tree carbon (mg/ha)")
```

```{r}

ggplot(plots_2017, aes(y= sample_count, x = sd_tree_c_ha)) +
  geom_point(aes(color = plot_size_m2))


```

